---
title: "GO Enrichment analysis for specific varroa modules"
author: "Nurit"
date: "2/5/2021"
output:
  html_document:
    theme: paper
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      df_print: paged
---

## Load libraries
```{r libraries, include=FALSE}
library("dplyr")
library("tidyverse")
library("vegan")
library("DESeq2")
library("ggfortify")
library("WGCNA")
library("rmarkdown")
library("knitr") # for the markdown
library("kableExtra") # for creating a scrolling table
library("ggplot2") # for plotting 
library("ape") # for reading the phylogenetic tree and for mantel.test
library("Biostrings")
library("ggtree") # for plotting the tree
library("ggrepel") # for spreading text labels on the plot
library("scales") # for axis labels notation
library("GO.db")
library("reshape2")
library("RSQLite")
library("AnnotationDbi")
library("GSEABase")
library("GOstats")
library("maps") # for the map background
library("leaflet") #for the interactive maps
library("htmltools")
library("rgdal")
library("grid")
library("gridExtra")
library("GeneOverlap")
library("cluster")
```

## seting working dir and loading data 
```{r eval=FALSE, include=FALSE}
#the working directory is in GitHb:
setwd("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks")

# Load the expression and trait data saved in the first part 
load(file = "/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/results/varroa_virus-01-dataInput.RData"); 

# Load network data saved in the second part.
load(file = "/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/results/Varroa_modules_networkConstruction-auto.RData")

annot.vd <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/data/VdesGOready2.csv") 
```

- data provided in [here](https://github.com/MaevaTecher/varroa-denovo-genomes/blob/master/data/Positive%20selection/Vdesselected1511.csv)
- codes provided in [here](https://maevatecher.github.io/varroa-denovo-genomes/#go_term_enrichment_of_positively_selected_genes)

## Goterm enrichment analysis
```{r eval=TRUE, echo=TRUE}
#Preparing the GO frame
annot.vd2 <- annot.vd %>%
  mutate(evidence = "IEA") %>%
  dplyr::select(go_id = GO.ids, evidence, gene = Gene.id)

head(annot.vd2)

goFrame.vd <-GOFrame(annot.vd2, organism = "Vd")
goAllFrame.vd <-GOAllFrame(goFrame.vd)
gsc.vd <-GeneSetCollection(goAllFrame.vd, setType = GOCollection())

#Preparing the universe
universe.vd <- unique(annot.vd2$gene)
head(universe.vd)

# Preparing the gene set (list of genes in an interesting module) 
# change "salmon" to the name of the desired module, in the first line: [moduleColors=="salmon"], and in the final "write.csv(file = "GO_term_enrichment_**salmon**BP.csv")
ME <- names(for_modules)[moduleColors=="black"]
ME_df <- data.frame(gene = ME)
genes.vd <- unique(ME_df$gene)
head(genes.vd)

params.vd <- GSEAGOHyperGParams(name = "Vd_GO_enrichment",
                                geneSetCollection = gsc.vd,
                                geneIds = genes.vd,
                                universeGeneIds = universe.vd,
                                ontology = "BP", # change with MF, CC to test all
                                pvalueCutoff = 0.05,
                                conditional = F,
                                testDirection = "over")

over.vd <- hyperGTest(params.vd)
over.vd
summary(over.vd)

GO_enrich.vd <- as.data.frame(summary(over.vd))
GO_enrich.vd %>% 
  arrange(Pvalue) %>% 
  write.csv(file = "/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_black.csv")
```

## plot the go-terms analysis, per module
```{r eval=F}
#load the go term enrichment analysis done using package ‘Category’
genesGO_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_salmon.csv")

# keep only highly significant terms
genesGO <- genesGO_0 %>%
  filter(Pvalue < 0.0001) 

#you can play with the number of terms by adjusting the "filter(Pvalue < 0.0000000001)"

#now plot it based on the count of each term and the pvalue:
g <- genesGO %>%
  mutate(logPv = log(Pvalue))

ggplot(g, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 13") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
  
```

## plot all significant modules go-terms
### module 13, salmon
```{r eval=F}
#load the go term enrichment analysis done using package ‘Category’
GOME_13_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_salmon.csv")

# keep only highly significant terms
GOME_13 <- GOME_13_0 %>%
  filter(Pvalue < 0.0001) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_13 <- ggplot(GOME_13, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 13") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```
 
### module 3, Brown:
```{r} 
#load the go term enrichment analysis done using package ‘Category’
GOME_3_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_brown.csv")

# keep only highly significant terms
GOME_3 <- GOME_3_0 %>%
  filter(Pvalue < 0.00001) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_3 <- ggplot(GOME_3, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 3") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```

### module 6, red:
```{r} 
#load the go term enrichment analysis done using package ‘Category’
GOME_6_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_red.csv")

# keep only highly significant terms
GOME_6 <- GOME_6_0 %>%
  filter(Pvalue < 0.01) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_6 <- ggplot(GOME_6, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 6") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```

### module 9, magenta:
```{r} 
#load the go term enrichment analysis done using package ‘Category’
GOME_9_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_magenta.csv")

# keep only highly significant terms
GOME_9 <- GOME_9_0 %>%
  filter(Pvalue < 0.0000000001) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_9 <- ggplot(GOME_9, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 9") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```
### module 1, turquoise:
```{r} 
#load the go term enrichment analysis done using package ‘Category’
GOME_1_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_turquoise.csv")

# keep only highly significant terms
GOME_1 <- GOME_1_0 %>%
  filter(Pvalue < 0.00001) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_1 <- ggplot(GOME_1, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 1") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```

### module 5, green:
```{r} 
#load the go term enrichment analysis done using package ‘Category’
GOME_5_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_green.csv")

# keep only highly significant terms
GOME_5 <- GOME_5_0 %>%
  filter(Pvalue < 0.001) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_5 <- ggplot(GOME_5, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 5") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```

### module 10, purple:
```{r} 
#load the go term enrichment analysis done using package ‘Category’
GOME_10_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_purple.csv")

# keep only highly significant terms
GOME_10 <- GOME_10_0 %>%
  filter(Pvalue < 0.00005) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_10 <- ggplot(GOME_10, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 10") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```

### module 14, cyan:
```{r} 
#load the go term enrichment analysis done using package ‘Category’
GOME_14_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_cyan.csv")

# keep only highly significant terms
GOME_14 <- GOME_14_0 %>%
  filter(Pvalue < 0.001) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_14 <- ggplot(GOME_14, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 14") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```

### module 7, black:
```{r} 
#load the go term enrichment analysis done using package ‘Category’
GOME_7_0 <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks/results/GO_term_black.csv")

# keep only highly significant terms
GOME_7 <- GOME_7_0 %>%
  filter(Pvalue < 0.001) %>%
  mutate(logPv = log(Pvalue))

#now plot it based on the count of each term and the pvalue:
GO_7 <- ggplot(GOME_7, aes(x = 0, y = Term, size = Count, color = logPv)) +
  scale_color_gradient(low= "#f0e11a", high="#f70a1a") +
  geom_point(alpha = 0.9) +
  geom_point(shape = 1,colour = "black") +
  theme_classic() +
  ggtitle("GOterm enrichment analysis \n module 7") +
  xlab(label = "") +
  ylab(label = "GOterm") +
  scale_x_discrete()
```
### plot all significant interacting modules:
```{r eval=FALSE, include=FALSE}
# modules positively interacting with ARV-2 and VOV_1 (module 3), and with VDV2 and VDV4 (modules 6, 9):
grid.arrange(GO_3, GO_6, GO_9,  nrow = 2, ncol =2)

# modules negatively interacting ONLY with DWVa: 
grid.arrange(GO_1 ,GO_5, GO_10, nrow = 1)

# modules positively interacting ONLY with DWVa: 
grid.arrange(GO_14, GO_7, nrow = 1)

# module negatively interacting with DWVa, DWVc, BQCV and VDV3
GO_13
```