---
title: "varroa_virus_networks"
author: "Sasha Mikheyev and Nurit Eliash"
date: "1/22/2020"
output:
  html_document:
    theme: paper
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      df_print: paged
---
this analysis is on the 66 libraries (i removed the 5 outlirered libraries from the "data/kallisto" folder in GitHub)
the working directory is in GitHb:
"/Users/nuriteliash/Documents/GitHub/varroa-virus-networks"

I save all the big outputs results in my local storage:
"/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/results"

# Load libraries
```{r load library, message=FALSE, warning=FALSE, echo=FALSE}
library("dplyr")
library("tidyverse")
library("vegan")
library("DESeq2")
library("ggfortify")
library("WGCNA")
library("rmarkdown")
library("knitr") # for the markdown
library("kableExtra") # for creating a scrolling table
library("ggplot2") # for plotting 
library("ape") # for reading the phylogenetic tree and for mantel.test
library("Biostrings")
library("ggtree") # for plotting the tree
library("ggrepel") # for spreading text labels on the plot
library("scales") # for axis labels notation
library("GO.db")
library("reshape2")
library("RSQLite")
library("AnnotationDbi")
library("GSEABase")
library("GOstats")
library("maps") # for the map background
library("leaflet") #for the interactive maps
library("htmltools")
library("rgdal")
library("grid")
library("gridExtra")
library("GeneOverlap")
library("cluster")
library("rmdformats")
library("corrplot")
library("viridis")
library("hrbrthemes")
library("ggthemes")
library("RColorBrewer")

knitr::opts_chunk$set(echo = TRUE)
```

# Gene network analysis
A gene co-expression network identify genes that interact with one another based on common expression profiles (Barabási and Oltvai 2004). Groups of co-expressed genes that have similar expression patterns across samples are identified using hierarchical clustering and are placed in gene ‘modules’ (Miller, Horvath, and Geschwind 2010). 
Weighted gene co-expression analysis was conducted using the WGCNA package in R (Langfelder and Horvath 2008).

## Data selection
We searched for “varroa” in the SRA databases (NCBI, January 2020) with the following filters:
 *  _Source_: "RNA"
 *  _Library Strategy_: “RNA-seq” (Random sequencing of whole transcriptome)
 *  _Library Source_: “TRANSCRIPTOMIC”, “VIRAL RNA”
 *  Only whole mites (not partial organs)

after filtering outlier libraries (based on PCA analysis), 
we used a total of 66 SRAs belonging to 11 studies of varroa RNAseq for the network analysis (details of the final libraries available in Supp table 1)

## Load data 
The input data frame "df" contains 35,669 isoforms of 66 varroa SRA libraries, downloaded from NCBI, using the 

consisted of a regularized log transformed matrix of 18,313 genes (as columns) and 29 libraries (as rows) of the venom gland which was filtered for low expressed transcripts (tpm > 0.05). Pm_7 was excluded due to low spike-in (see methods)

mapping RNAseq reads to viruses and varroa genes
```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
setwd("/Users/nuriteliash/Documents/GitHub/varroa-virus-networks")

# read the viruses IDs:
# awk '$3=="mRNA"' GCF_002443255.1_Vdes_3.0_genomic.gff | perl -n -e 'print "$1\t$2\n" if /gene=LOC(\d+);.*transcript_id=(.*)/' | gzip >! gene2isoform.txt.gz
virusId <- read_tsv("data/viruses.txt", col_names = c("target_id", "description"))
# rename all 6 VOV_1 segemnts with the same name: "VOV_1"
virusId[3:8, 2] <- "VOV_1"

# map the Varroa and viruses' reads
# first make a function for mapping the reads
read_kallisto <- function(filename) {
  sampleName <- sub("data/kallisto/(.*).tsv.gz","\\1", filename)
  return(read_tsv(filename) %>%
           dplyr::select(!!sampleName := tpm))
}

# map the reads in the SRA libraries to varroa and viruses genomes 
df <- list.files(path = "data/kallisto", full.names = TRUE) %>% 
  lapply(read_kallisto) %>% 
  bind_cols() 

# add a column "target_id" with the genes id
df$target_id <- list.files(path = "data/kallisto", full.names = TRUE)[1] %>% 
  read_tsv() %>% 
  dplyr::select(target_id) %>% 
  pull()

# save the virus list and the isoform tpm as RDS 
saveRDS(df, file = "/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/data/kallisto.rds")
saveRDS(virusId, file = "/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/data/virusId.rds")
```

## AIM: describing viruses' abundance in the different varroa SRA libraries
first make the suitable data frame containing virus load in each library
```{r first make the suitable data frame containing virus load in each library, message=FALSE, warning=FALSE}
x <- left_join(virusId, df, by = "target_id") %>% dplyr::select(-target_id) 
  
# VOV_1. virus genome has been described in 6 different segments (Levin et al. 2016). 
# so we sum the 6 segments tpms in each library in a separate table 'VOV_1':
    VOV_1 <- x %>% 
      dplyr::slice(3:8) %>%
      group_by(description) %>%
      summarize_all(sum)
#filter out the segments of VOV_1,  
x <- dplyr::filter(x, description != "VOV_1")

# and bind the two data frames together 
viruses_load <- rbind(x, VOV_1) 

# reorder the viruses, so VOV_1 will appear in row number 3
r <- 2
insertRow <- function(x, VOV_1, r) {
  x[seq(r+1,nrow(x)+1),] <- x[seq(r,nrow(x)),]
  x[r,] <- VOV_1
  x
}

insertRow(x, VOV_1, r)
head(viruses_load)
saveRDS(viruses_load, file = "/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/results/viruses_load.RData")
```
 
### vizualize viruses distribution and abundance in (1)abundance plot; (2)PCA of viruses; (3)heat-map
```{r message=FALSE, warning=FALSE}
df <- readRDS("/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/data/kallisto.rds")
viruses_load <- readRDS(file = "/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/results/viruses_load.RData")

# (1) virus abundance plot
joined_virus_lib <- viruses_load %>% gather("library", "tpm", -description)

ggplot(joined_virus_lib, aes(x= library, y = tpm, fill = description)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90)) 

# (2) plot a PCA of viruses abundance 
joined_virus_lib_PCA <- joined_virus_lib %>% group_by(description, library) %>% summarise(tpm = sum(tpm)) %>% spread(description, tpm) %>% dplyr::select(-library) 

virus_PCA <- princomp(joined_virus_lib_PCA)  

biplot(virus_PCA)

# (3) heat-map ov viruses loads (log tpm) per library
# plot the viruses abundance in each library

# re-order the viruses, based on viral load (from high to low):
#total_virus_load <- viruses_load %>%
 # gather("library", "tpm", -description) %>% 
  #group_by(description) %>%
  #summarise(total_load = sum(tpm)) %>%
  #arrange(desc(total_load))
# lock in factor level order of the total viral load
#total_virus_load$description <- factor(total_virus_load$description, levels = total_virus_load$description)
# make a vector of the new viruses order

new_order <- c("DWVa", "VDV1/DWVb", "VDV2", "BMV","VOV_1","ARV_2" , "DWVc", "AmFV", "VDV3", "ABPV", "VTLV", "SV", "BQCV","VDV4", "IAPV", "KBV", "SBPV", "LSV","CBPV" ,"AFV", "ANV", "VPVL_46","VPVL_36")

# re-order the viruses, based on viral load and viral abundance:
viruses_load_arranged <- viruses_load  %>%
  mutate(description =  factor(description, levels = new_order)) %>%
  arrange(description) 
  
# lock in factor level order
viruses_load_arranged$description <- factor(viruses_load_arranged$description, levels = viruses_load_arranged$description)

# spread the df with tpm in logscale
virus_abund <- viruses_load_arranged %>%  
    gather("library", "tpm", -description) 

virus_abund[, 3] <- (virus_abund[,3]+ 0.000001)
virus_abund[, 3] <- log10(virus_abund[,3]) 

# now make a heat map
ggplot(data = virus_abund, mapping = aes(x = description,
                                                       y = library,
                                                       fill = tpm)) +
  scale_fill_gradient2(low="#FFFFCC", mid = "#FF9933", high="#990000", space = "Lab") +
  geom_tile() +
  theme_linedraw() +
  xlab(label = "Viruses") +
  ylab(label = "Varroa SRA libraries") +
  labs(fill = "Viral load \n log(tpm)")  + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  + 
  theme(axis.ticks = element_blank())


  

# or use heatmap() function
x <- viruses_load %>%
  column_to_rownames("description") %>%
  as.matrix()
heatmap(x, Colv = NA, xlab="Libraries", ylab="Viruses", cexRow=1, labRow=paste(rownames(x)))


#FFFFCC - cream color
#scale= "row",
#Rowv = NA, 
# virus_heatmap + scale_color_brewer(palette = "BuPu")

#scale_fill_manual(values=c("red", "blue", "green")) +
#scale_fill_gradient(low="cream", high="blue", space = "Lab")
 #theme(axis.text.x = element_text(angle = 30)) 
 # scale_color_brewer(palette = "RdYlBu") 
#coul <- colorRampPalette(brewer.pal(8, "Blues"))(25)
#scale_fill_gradient2(low="#FFFFCC", mid = "#FF9933", high="#990000", space = "Lab")
```

## virus Correlation matrix: 66 libs
```{r prepare virus data: 66, message=FALSE, warning=FALSE}
# plot the viruses abundance in each library 

# prepare the 'viruses_load' table, with 17 viruses, 
# excluding viruses with 'zero tpm': "CBPV", "AFV", "ANV" , "VPVL_46", "VPVL_36", "SBPV", also excluding viruses with low abundance :"KBV", "LSV".  left with 15 viruses
a_viruses_load_66 <- viruses_load %>%
  filter(description %in% c("DWVa", "DWVc", "IAPV",  "ABPV", "BQCV","SV", "VDV1/DWVb", "VDV2", "VDV3", "BMV", "ARV_2", "AmFV","VTLV","VDV4","VOV_1")) %>%
  column_to_rownames("description") %>% 
  transposeBigData() %>%
  rownames_to_column("library") %>%
  #identify and remove the row numbers of the outlierd libraries (found by PCA) 
  #filter(!(library %in% c("SRR5109825", "SRR5109827", "SRR533974" , "SRR3927496", "SRR8867385"))) %>% 
  column_to_rownames("library")

# correlate viruses load (66 libs)
corrplot2 <- function(data,
                      method = "pearson",
                      sig.level = 0.05,
                      order = "original",
                      diag = FALSE,
                      type = "upper",
                      tl.srt = 90,
                      number.font = 1,
                      number.cex = 0.5,
                      mar = c(0, 0, 0, 0)) {
  
  data_incomplete <- data
  data <- data[complete.cases(data), ]
  mat <- cor(data, method = method)
  cor.mtest <- function(mat, method) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat <- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
      for (j in (i + 1):n) {
        tmp <- cor.test(mat[, i], mat[, j], method = method)
        p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
      }
    }
    colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
    p.mat
  }
  p.mat <- cor.mtest(data, method = method)
  col <- colorRampPalette(c("#FFFFCC", "#FF9933","#990000"))
  corrplot(mat,
    method = "color", col = col(200), number.font = number.font,
    mar = mar, number.cex = number.cex,
    type = type, order = order,
    addCoef.col = "black", # add correlation coefficient
    tl.col = "black", tl.srt = tl.srt, # rotation of text labels
    # combine with significance level
    p.mat = p.mat, sig.level = sig.level, insig = "blank",
    # hide correlation coefficients on the diagonal
    diag = diag
  )
}

virusAbundCor_66 <- corrplot2(
  data = a_viruses_load_66,
  method = "pearson",
  sig.level = 0.05,
  order = "original",
  diag = FALSE,
  type = "upper",
  tl.srt = 75
)

# hcl.colors(n = 50, palette = "RdYlBu", alpha = NULL, rev = FALSE, fixup = TRUE)
# col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

# save the matrice for future analysis (Mantel)
save(virusAbundCor_66, file = "/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/results/virusAbundCor_66.RData")
```

## virus Correlation matrix: 57 libs
```{r prepare virus data: 57, message=FALSE, warning=FALSE}
# make your data set:
# plot the viruses abundance in each library 

# prepare the 'viruses_load' table, with 17 viruses, 
# excluding viruses with 'zero tpm': "CBPV", "AFV", "ANV" , "VPVL_46", "VPVL_36", "SBPV", also excluding viruses with low abundance :"KBV", "LSV".  left with 15 viruses
a_viruses_load_57 <- viruses_load %>%
  filter(description %in% c("DWVa", "DWVc", "IAPV",  "ABPV", "BQCV","SV", "VDV1/DWVb", "VDV2", "VDV3", "BMV", "ARV_2", "AmFV","VTLV","VDV4","VOV_1")) %>%
  column_to_rownames("description") %>% 
  transposeBigData() %>%
  rownames_to_column("library") %>%
  #identify and remove the row numbers of the outlierd libraries (found by PCA) 
  #filter(!(library %in% c("SRR5109825", "SRR5109827", "SRR533974" , "SRR3927496", "SRR8867385"))) %>% 
  # remove 9 male libraries:
  filter(!(library %in% c("SRR5760848", "SRR5760838", "SRR5760828", "SRR5760818", "SRR5377265" ,"SRR5377270" ,"SRR5377269" ,"SRR5377263" ,"SRR5377266"))) %>%
  column_to_rownames("library")

# correlate viruses load-57 libs
corrplot2 <- function(data,
                      method = "pearson",
                      sig.level = 0.05,
                      order = "original",
                      diag = FALSE,
                      type = "upper",
                      tl.srt = 90,
                      number.font = 1,
                      number.cex = 0.5,
                      mar = c(0, 0, 0, 0)) {
  
  data_incomplete <- data
  data <- data[complete.cases(data), ]
  mat <- cor(data, method = method)
  cor.mtest <- function(mat, method) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat <- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
      for (j in (i + 1):n) {
        tmp <- cor.test(mat[, i], mat[, j], method = method)
        p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
      }
    }
    colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
    p.mat
  }
  p.mat <- cor.mtest(data, method = method)
  col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
  corrplot(mat,
    method = "color", col = col(200), number.font = number.font,
    mar = mar, number.cex = number.cex,
    type = type, order = order,
    addCoef.col = "black", # add correlation coefficient
    tl.col = "black", tl.srt = tl.srt, # rotation of text labels
    # combine with significance level
    p.mat = p.mat, sig.level = sig.level, insig = "blank",
    # hide correlation coefficiens on the diagonal
    diag = diag
  )
}

virusAbundCor_57 <- corrplot2(
  data = a_viruses_load_57,
  method = "pearson",
  sig.level = 0.05,
  order = "original",
  diag = FALSE,
  type = "upper",
  tl.srt = 75
)

# save the matrice for future analysis (Mantel)
save(virusAbundCor_57, file = "/Users/nuriteliash/OneDrive - OIST/Repos/varroa-virus-networks-Local/results/virusAbundCor_57.RData")
```

